R Notebook for MSDepression Project

### 2022-09-16 lab_notebook/lab_notebook.Rmd
Add files via upload

[] make figure of depression network from bold
  [x] resampled the harvard map to mni_icbm_2009b template: 
  [x] multiplied the resampled image with the mask
  [x] displayed in fsleyes and looks amazing
    underlay: mni_icbm152_t1_tal_nlin_asym_09a_mask.nii (drop intensity a bit to get the background not to have that ring around it)
    overlay: DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii (set to 3.09/3.09 for yellow, 3.09/5 for colored)

@@

>>>>

[x] make figure of depression network from bold
  [x] resampled the harvard map to mni_icbm_2009b template: 
  [x] multiplied the resampled image with the mask
  [x] displayed in fsleyes and looks amazing
    underlay: mni_icbm152_t1_tal_nlin_asym_09a_mask.nii (drop intensity a bit to get the background not to have that ring around it)
    overlay: DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii (set to 3.09/3.09 for yellow, 3.09/5 for colored)

@@

#### 20220913
1. [o] Flow chart (text is small, do I need to make bigger?)
  efile a.[x] demographics full
  efile b.[x] (promis scores)
  efile c. [] pipeline is e -file (can only have max 4 pt figure), will need to work with Maddie for this.
2. Making depression network
  [x]: Cortical rendering of depression network
  [x]: DSI fascicle of volume of overlap w/network (how to scale for visualizations) (new hampshire forest in the fall colors)
3.  In vs outside dep net
    a. inside vs outside dep net violin
    b. pink and blue fascicle picture part B
4. Diagnosis analysis
    a. no diff by lesion volume (violin)
    b. but difference in fascicle volume (violin)
        - just in/outside of network (put stuff about within/outside network for each diagnostic group in text)
5. Dep Network/dx interactions
     a. interaction graph
     b. scatter (effect size by overlap)
     c. Fascicles colored by effect size (edited)

#### 20220914
Ted meeting

[] Elizabeth is currently replicating
  - found a bug in the naming order of one of the files. We have changed it, and the subsequent results are the same, but a little better
  - The mask stays the same (dep vs non), so no need to regenerate those pictures
  - Coloring for new hamphire colors and pink effect size are slightly different
    - [x] redid the new hampshire color plot
    - [x] redid pink plot
  - Will need to remake the main effect of dx and network violins with new data
    [x] main effect of network
    [x] main effect of diagnosis
  - Remake scatter and interaction plots
    [x] scatter
    [x] interaction
[x] review some sample figures
[x] talk about science for LRP
  - spoke with Mark, feasibility important
  - In line with that, do I convert aim 1 to 2 aims (have sex differences be one)?
  - Or (my preference), basically keep everything the same but state that aim 2 will be the beginning of a collection of data (50 people per year), rather than 250. 
    - would be much eaasier to do this rather than rewrite that whole chunk.
    - Also, mentorship team from K, but without mike fox and tom. Substitute sex differences for EMR training?
  - I have also included that I applied for a K already, and what my plans are if I don't get it
  -

#### 20220915
  [] cleaning up scatter with Elizabeth - incremented the volume by 1, and log transformed for a much prettier graph!
  [x] Making the binary brain map on an inflated brain in workbench
   - first, binarized it 
```{bash}
3dcalc -a DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked.nii -expr 'ispositive(a-3.09)' -prefix DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii
```
   - tried to use command line tools for workbench, couldn't do it, but val passed on a script that was just amazing, and I adapted it for my stuff
```{r}

---
title: "Project an MNI volume to the fslr surface for visualization"
author: "Valerie Jill Sydnor, adapted by Erica Baller"
date: "9/16/2022"
output: html_document
---
library(ciftiTools)
ciftiTools.setOption('wb_path', '/Users/eballer/Software/workbench/')
#set the path to your installation of connectome workbench as 'wb_path'
library(rgl) #to use ciftiTools graphics
library(rmarkdown)
library(gifti) #to read in your surface giftis

#Map volume data to left and right cortical surfaces
#left hemisphere
command1=sprintf("-volume-to-surface-mapping /Users/eballer/BBL/msdepression/templates/harvard_depression/surfaces/DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_l.shape.gii -trilinear")
ciftiTools::run_wb_cmd(command1, intern = FALSE, ignore.stdout = NULL, ignore.stderr = NULL)

#right hemisphere
command2=sprintf("-volume-to-surface-mapping /Users/eballer/BBL/msdepression/templates/harvard_depression/surfaces/DepressionCircuit_t_resampled_to_mni_icbm152_asym09a_masked_binarized_3_09.nii /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_r.shape.gii -trilinear")
ciftiTools::run_wb_cmd(command2, intern = FALSE, ignore.stdout = NULL, ignore.stderr = NULL)

#Read in your surface giftis
test_l <- read_gifti("/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_l.shape.gii")
test_r <- read_gifti("/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf_r.shape.gii")

#Combine your giftis into one cifti surface (for visualization with ciftiTools)
test.ciftimap <- as_cifti(cortexL = test_l$data$normal, cortexR = test_r$data$normal) #the datas are stored in file$data$normal
test.ciftimap$data$cortex_left[test.ciftimap$data$cortex_left > 0] <- 1
test.ciftimap$data$cortex_right[test.ciftimap$data$cortex_right > 0] <- 1

#You can also save out this cifti for viewing in connectome workbench
write_cifti(test.ciftimap, "/Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf.dscalar.nii")

#Choose what surface you want to visualize your data on (midthickness, inflated, sphere, etc.)
#lets use midthickness
surfL_name <- read_surf("/Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii")
surfR_name <- read_surf("/Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii")

#gotta add these surfs to your cifti from the step above
test.ciftimap <- add_surf(test.ciftimap, surfL=surfL_name, surfR=surfR_name)

#Visualize!
#```{r, rgl=TRUE, format="png"} 
view_cifti(test.ciftimap, colors=c("magma"), widget=TRUE)
```

- I then had to take this into workbench, and futz with the colors

  - open application: wb_view
  - open file: /Users/eballer/BBL/msdepression/templates/harvard_depression/testvol2surf.dscalar.nii
  - open file: /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.L.midthickness_MSMAll.32k_fs_LR.surf.gii
  - open file: /Users/eballer/Software/workbench/workbench_files/HCP_S1200_GroupAvg_v1/S1200.R.midthickness_MSMAll.32k_fs_LR.surf.gii
  - click montage
  - under settings, click the wrench
  - Threshold: Off
  - Palette
    - Pick the palette HSB8_clrmid
    - Invert off
    - click interpolate colors
    - click fixed
    - setting 40/0/0/0
  - click negative and positive
  - Click bars under histogram
  - change bars color and envelope to fuchsia (don't know if this does anything)
  - Color bar and palette settings: apply to all maps

- Save, screenshot, bring to illustrator and have at it

- Also, will do log scatter, looks so much better. Have to increment by 1 in order to make it pretty though, otherwise all the 0 values will screw you over
